Introducing ICBe: Very High Recall and Precision Event Extraction from Narratives about International Crises

### 2023-05-30 replication_paper/psrm_draft/ICBe_manuscript.Rmd
start clarifying new elements of ontology

This measurement paper makes the following arguments. First, there is a real-world unobserved latent concept known as international relations that can be systematically measured. Second, we propose a method for systematic large-scale measurement of the actors and behaviors in international affairs and as a proof of concept apply that method to a well-regarded and salient sample of international events known as international crises. Third, in doing so we confirm that those measurements exhibit several desirable kinds of internal and external validity and out-perform existing approaches. The article organizes that argument into sections. A final section concludes.

@@

The columns compare the ontological coverage of ICBe to existing state of the art systems in production and with global coverage. They begin with our contribution ICBe, alongside other event-level datasets including CAMEO dictionary lookup-based systems (Historical Phoenix [@althausClineCenterHistorical2019]; ICEWS [@boscheeICEWSCodedEvent2015]; Terrier [@grantOUEventData2017]), the Militarized Interstate Disputes Incidents dataset, and the UCDP-GED dataset [@daviesOrganizedViolence19892022; @sundbergIntroducingUCDPGeoreferenced2013].[^3] The final set of columns compares episode-level datasets beginning with the original ICB project [@brecherInternationalCrisisBehavior2021; @brecherCrisesWorldPolitics1982; @beardsleyInternationalCrisisBehavior2020]; the Militarized Interstate Disputes dataset [@palmerMID5Dataset20112021; @giblerInternationalConflicts181620102018], and the Correlates of War [@sarkeesResortWar181620072010]. With the exception of large-scale CAMEO dictionary-based systems, the existing state of the art quantitative datasets ignore the vast majority of the information content found in international relations.[^4]

@@

[^4]: See @balaliCOfEEComprehensiveOntology2021 for a recent review of ontological depth and availability of Gold Standard example text.

The rows represent the types of information we expect to find in IR and forms the basis for our proposed ontology. We employ a metaphor of a chess game, with players (polities, rebel groups, IGOs, etc.), pieces (military platforms, civilians, domains), and behaviors (think, say, do). Precise sequencing is required to capture gambits (sequences of moves) and outcomes (victory, defeat, peace, etc.), while precise geo-coding is required to understand the chessboard (medium of conflict). We find `r codings_long_agreement %>% filter(varname_normalized %>% str_detect("actor")) %>% pull(value_normalized) %>% unique() %>% length()` actors and 117 different behaviors.[^5]

[^5]: See the full codebook on Github Repository [ICBEventData](https://urldefense.com/v3/__https://github.com/CenterForPeaceAndSecurityStudies/ICBEventData__;!!Mih3wA!WxDJtEczKfxGTh0S2Krunap8ReymFEL5iTWaSfOHeqlSdyfRx77zmjBSWO1OAm13$).

@@

[^8]: As the ICB narratives and ICBe are intended to explain the conflictual behavior in a political context, many of the missing events concern the economic components of conflict events (eg. nationalizing a foreign business). ICBe's limitation at this stage is its focus on international events that involve conflict (ICB crises), but even with this limitation in the information we trained on ICBe is more comprehensive than the existing datasets that *do* try to code the economic dimensions of these conflicts. Since expanding the ontology to broader international phenomenon is an easier problem to address than the over-inclusion of irrelevant information, we see this as promising for future implementation of our model.

>>>>

This measurement paper makes the following arguments. First, there is a real-world unobserved latent concept known as international relations that can be systematically measured. Second, we propose a method for systematic large-scale measurement of the actors and behaviors in international affairs and as a proof of concept apply that method to a well-regarded and salient sample of international events known as international crises. Third, in doing so we confirm that those measurements exhibit several desirable kinds of internal and external validity and out-perform existing approaches. The article organizes that argument into XX sections. A final section concludes.

@@

The columns compare the ontological coverage of ICBe to existing state of the art systems in production and with global coverage. They begin with our contribution ICBe, alongside other event-level datasets including CAMEO dictionary lookup-based systems (Historical Phoenix [@althausClineCenterHistorical2019]; ICEWS [@boscheeICEWSCodedEvent2015]; Terrier [@grantOUEventData2017]), the Militarized Interstate Disputes Incidents dataset, and the UCDP-GED dataset [@daviesOrganizedViolence19892022; @sundbergIntroducingUCDPGeoreferenced2013].[^3] The final set of columns compares episode-level datasets beginning with the original ICB project [@brecherInternationalCrisisBehavior2021; @brecherCrisesWorldPolitics1982; @beardsleyInternationalCrisisBehavior2020]; the Militarized Interstate Disputes dataset [@palmerMID5Dataset20112021; @giblerInternationalConflicts181620102018], and the Correlates of War [@sarkeesResortWar181620072010].

@@

The rows represent the types of information we expect to find in IR and forms the basis for our proposed ontology. We employ a metaphor of a chess game, with players (polities, rebel groups, IGOs, etc.), pieces (military platforms, civilians, domains), and behaviors (think, say, do). Precise sequencing is required to capture gambits (sequences of moves) and outcomes (victory, defeat, peace, etc.), while precise geo-coding is required to understand the chessboard (medium of conflict). We find `r codings_long_agreement %>% filter(varname_normalized %>% str_detect("actor")) %>% pull(value_normalized) %>% unique() %>% length()` actors and 117 different behaviors.[^4]

[^4]: See the full codebook on Github Repository [ICBEventData](https://urldefense.com/v3/__https://github.com/CenterForPeaceAndSecurityStudies/ICBEventData__;!!Mih3wA!WxDJtEczKfxGTh0S2Krunap8ReymFEL5iTWaSfOHeqlSdyfRx77zmjBSWO1OAm13$).

Table 1 demonstrates that with the exception of large-scale CAMEO dictionary-based systems, our ontology improves upon the the existing state of the art quantitative datasets that ignore important information content about international interactions.[^5] In particular,

[^5]: See @balaliCOfEEComprehensiveOntology2021 for a recent review of ontological depth and availability of Gold Standard example text.

-   separating chess pieces from players (military vs civilian)

-   addition of domains

@@

[^8]: As the ICB narratives are intended to explain conflictual behavior in a political context, many of the missing events concern more economic components of conflict (eg. nationalizing a foreign business). Even when they occur in the context of a crisis, these events largely fall outside the sample of information on which ICBe's ontology is currently trained. Even with this limitation, ICBe is more comprehensive than the existing datasets that *do* try to code the economic dimensions of these crises. We see expanding the ontology to broader international phenomenon as a promising future implementation of our model.
